On the adhesivity of EGGs

### 2025-03-02 calco/main.tex
pulizia

\begin{definition}\label[definition]{def:tg}\index{term graph}
	Given an algebraic signature $\Sigma$, we say that a labelled hypergraph $(l, !_{V_\mathcal{G}})\colon \mathcal{G}\to \mathcal{G}^{\Sigma}$ is a \emph{term graph} if $t_\mathcal{G}$ is mono. We define $\tg$ to be the full subcategory of $\hyp_{\Sigma}$ and denote by $I_\Sigma$ the inclusion. Restricting $U_\Sigma\colon \hyp_{\Sigma}\to \catname{Set}$ we get a forgetful functor $U_{\tg}\colon \tg\to \catname{Set}$.
\end{definition}

\begin{remark}\label[remark]{rem:mono2}By \Cref{rem:label}, we know that if $\mathcal{G}$ is a term graph then $t_{\mathcal{G}}=v_1\circ \tau_{\mathcal{G}}$, where $v_1$ is the coprojection of $V_{\mathcal{G}}$ into $V^\star_{\mathcal{G}}$.  Notice that since $t_{\mathcal{G}}$ is mono then $\tau_{\mathcal{G}}$ is mono too.
\end{remark}

@@

\begin{corollary}\label{cor:mono1}
	An arrow $h = (h_E, h_V, h_C): \mathcal{G \to H}$ in $\EqHyp$ is mono if and only if $(h_E, h_V)$ is a mono in $\hyp$.
\end{corollary}

@@

\todo{Ripensandoci il funtore quoziente non serve: l'unico punto in cui si usa nella tesi è che il quoziente di un colimite è il colimite dei quozienti, ma questo segue dal punto $1$ di \Cref{prop:eqhyp_complete}. Inoltre la dimostrazione della creazione è sbagliata (e probabilmente quella è una cosa falsa), io cancellerei}

@@

On the one hand, if we apply $\quo$ and compute the pullback of the resulting morphisms, we obtain an hypergraph with no hyperdeges and $A\times B$ as set of nodes. On the other hand, \Cref{prop:eqhyp_complete} we know that taking the pullback of the original two morphisms gives us the hypergraph with equivalence $(\emptyset, \emptyset, Q, ?_{\emptyset^\star}, ?_{\emptyset^\star}, q)$ where $Q$ is the image of $?_{A\times B}$, so that $Q$ is also empty.  
\end{example}

@@

\begin{proof}
	Let $\mathcal{G}_i = (A_i, B_i, Q_i, s_i, t_i, q_i)$, $\mathcal{G}'_i=(A'_i, B'_i, Q'_i, s'_i, t'_i, q'_i)$, for $i \in \{1, 2, 3, 4\}$, be hypergraphs with equivalence, 
	and that suppose in the first diagram below,  all the vertical faces are pullbacks, the bottom face is a pushout and $h$ is regular mono. By \Cref{prop:eqhyp_complete} the same is true for the other two cubes, by \Cref{cor:mono2}, $h_E$ and $h_V$ are mono and so the top faces of these cubes are pushouts.
	\[
	\xymatrix@C=10pt@R=10pt{&\mathcal{G}_1'\ar[dd]|\hole_(.65){a}\ar[rr]^{h'} \ar[dl]_{k'} && \mathcal{G}_2' \ar[dd]^{b} \ar[dl]_{z'} & &A_1'\ar[dd]|\hole_(.65){a_E}\ar@{>->}[rr]^{h_E'} \ar[dl]_{k_E'} && A_2' \ar[dd]_{b_E} \ar[dl]_{z_E'} &&B_1'\ar[dd]|\hole_(.65){a_V}\ar@{>->}[rr]^{h_V'} \ar[dl]_{k_V'} && B_2' \ar[dd]^{b_V} \ar[dl]_{z_V'} \\ 
		\mathcal{G}_3'  \ar[dd]_{c}\ar[rr]^(.7){p'} & & \mathcal{G}_4' \ar[dd]_(.3){d} &&A_3'  \ar[dd]_{c_E}\ar@{>->}[rr]^(.7){p_E'} & & A_4' \ar[dd]_(.3){d_E}
		&&B_3'  \ar[dd]_{c_V}\ar@{>->}[rr]^(.7){p_V'} & & B_4' \ar[dd]_(.3){d_V}\\
		&\mathcal{G}_1\ar[rr]|\hole^(.65){h} \ar[dl]^{k} && \mathcal{G}_2 \ar[dl]^{z} & &A_1\ar@{>->}[rr]|\hole^(.65){h_E} \ar[dl]^{k_E} && A_2 \ar[dl]^{z_E} &&B_1\ar@{>->}[rr]|\hole^(.65){h_V} \ar[dl]^{k_V} && B_2 \ar[dl]^{z_V} \\
		\mathcal{G}_3 \ar[rr]_{p} & & \mathcal{G}_4 && A_3 \ar@{>->}[rr]_{p_E} & & A_4&&B_3 \ar@{>->}[rr]_{p_V} & & B_4 }
	\]

@@

\begin{proposition}\label{prop:monos_in_eqhyps}
	A morphism $h$ between two objects of $\EqHyps$ is mono if and only if $T(h)$ is mono.
\end{proposition}

@@

\quot*
\begin{proof}\label{proof:quot}
  Given an hypergraph $\mathcal{H}:=(E, V, s, t)$, let us define $R(E, V, s, t)$ as $(E, V, V, s, t, \id{B})$. By construction $\quo(R(\mathcal{{H}})) =\mathcal{H}$ so that we can take the identity as $\epsilon_{\mathcal{H}}\colon \quo(R(\mathcal{H}))\to \mathcal{H} $.

@@

By the proof of the first point of \Cref{prop:eqhyp_complete} the . \qedhere }

>>>>

\begin{definition}\label[definition]{def:tg}\index{term graph}
	Given an algebraic signature $\Sigma$, we say that a labelled hypergraph $(l, !_{V_\mathcal{G}})\colon \mathcal{G}\to \mathcal{G}^{\Sigma}$ is a \emph{term graph} if $t_\mathcal{G}$ is a mono. We define $\tg$ to be the full subcategory of $\hyp_{\Sigma}$ and denote by $I_\Sigma$ the inclusion. Restricting $U_\Sigma\colon \hyp_{\Sigma}\to \catname{Set}$ we get a forgetful functor $U_{\tg}\colon \tg\to \catname{Set}$.
\end{definition}

\begin{remark}\label[remark]{rem:mono2}By \Cref{rem:label}, we know that if $\mathcal{G}$ is a term graph then $t_{\mathcal{G}}=v_1\circ \tau_{\mathcal{G}}$, where $v_1$ is the coprojection of $V_{\mathcal{G}}$ into $V^\star_{\mathcal{G}}$.  Notice that since $t_{\mathcal{G}}$ is a mono then $\tau_{\mathcal{G}}$ is a mono too.
\end{remark}

@@

\begin{corollary}\label{cor:mono1}
	An arrow $h = (h_E, h_V, h_C): \mathcal{G \to H}$ in $\EqHyp$ is a mono if and only if $(h_E, h_V)$ is a mono in $\hyp$.
\end{corollary}

@@

\commentato{
\todo{Ripensandoci il funtore quoziente non serve: l'unico punto in cui si usa nella tesi è che il quoziente di un colimite è il colimite dei quozienti, ma questo segue dal punto $1$ di \Cref{prop:eqhyp_complete}. Inoltre la dimostrazione della creazione è sbagliata (e probabilmente quella è una cosa falsa), io cancellerei}

@@

On the one hand, if we apply $\quo$ and compute the pullback of the resulting morphisms, we obtain an hypergraph with no hyperdeges and $A\times B$ as set of nodes. On the other hand, \Cref{prop:eqhyp_complete} we know that taking the pullback of the original two morphisms gives us the hypergraph with equivalence $(\emptyset, \emptyset, Q, ?_{\emptyset^\star}, ?_{\emptyset^\star}, q)$ where $Q$ is the image of $?_{A\times B}$, so that $Q$ is also empty.  
\end{example}
}

@@

\begin{proof}
	Let $\mathcal{G}_i = (A_i, B_i, Q_i, s_i, t_i, q_i)$, $\mathcal{G}'_i=(A'_i, B'_i, Q'_i, s'_i, t'_i, q'_i)$, for $i \in \{1, 2, 3, 4\}$, be hypergraphs with equivalence, 
	and that suppose in the first diagram below,  all the vertical faces are pullbacks, the bottom face is a pushout and $h$ is a regular mono. By \Cref{prop:eqhyp_complete} the same is true for the other two cubes, by \Cref{cor:mono2}, $h_E$ and $h_V$ are monos and so the top faces of these cubes are pushouts.
	\[
	\xymatrix@C=10pt@R=10pt{&\mathcal{G}_1'\ar[dd]|\hole_(.65){a}\ar[rr]^{h'} \ar[dl]_{k'} && \mathcal{G}_2' \ar[dd]^{b} \ar[dl]_{z'} & &A_1'\ar[dd]|\hole_(.65){a_E}\ar@{>->}[rr]^{h_E'} \ar[dl]_{k_E'} && A_2' \ar[dd]_{b_E} \ar[dl]_{z_E'} &&B_1'\ar[dd]|\hole_(.65){a_V}\ar@{>->}[rr]^{h_V'} \ar[dl]_{k_V'} && B_2' \ar[dd]^{b_V} \ar[dl]_{z_V'} \\ 
		\mathcal{G}_3'  \ar[dd]_{c}\ar[rr]^(.7){p'} & & \mathcal{G}_4' \ar[dd]_(.3){d} &&A_3'  \ar[dd]_{c_E}\ar@{>->}[rr]^(.7){p_E'} & & A_4' \ar[dd]_(.3){d_E}
		&&B_3'  \ar[dd]_{c_V}\ar@{>->}[rr]^(.7){p_V'} & & B_4' \ar[dd]_(.3){d_V}\\
		&\mathcal{G}_1\ar[rr]|\hole^(.65){h} \ar[dl]^{k} && \mathcal{G}_2 \ar[dl]^{z} & &A_1\ar@{>->}[rr]|\hole^(.65){h_E} \ar[dl]^{k_E} && A_2 \ar[dl]^{z_E} &&B_1\ar@{>->}[rr]|\hole^(.65){h_V} \ar[dl]^{k_V} && B_2 \ar[dl]^{z_V} \\
		\mathcal{G}_3 \ar[rr]_{p} & & \mathcal{G}_4 && A_3 \ar@{>->}[rr]_{p_E} & & A_4&&B_3 \ar@{>->}[rr]_{p_V} & & B_4 }
	\]

@@

\begin{proposition}\label{prop:monos_in_eqhyps}
	A morphism $h$ between two objects of $\EqHyps$ is a mono if and only if $T(h)$ is a mono.
\end{proposition}

@@

\commentato{
\quot*
\begin{proof}\label{proof:quot}
  Given an hypergraph $\mathcal{H}:=(E, V, s, t)$, let us define $R(E, V, s, t)$ as $(E, V, V, s, t, \id{B})$. By construction $\quo(R(\mathcal{{H}})) =\mathcal{H}$ so that we can take the identity as $\epsilon_{\mathcal{H}}\colon \quo(R(\mathcal{H}))\to \mathcal{H} $.

@@

By the proof of the first point of \Cref{prop:eqhyp_complete} the . \qedhere }
}
